Labor Markets and Technological Change: Evidence from Electronic Health Records

### 2021-10-28 Presentations/ThirdYearPaperpres.tex
Updates from Overleaf

\setbeameroption{show notes on second screen=right}
\setbeamertemplate{note page}{\pagecolor{yellow!5}\insertnote}\usepackage{palatino}
\usepackage{xspace}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}

@@

\begin{frame}[fragile]{What is an EHR?}
\begin{itemize}
    \item Technology purchased and implemented by hospitals
\end{itemize}

\begin{itemize}
    \item Used for:
    \vspace{3mm}
    \begin{itemize}
        \item Electronic patient records
        \vspace{3mm}
        \item More complex (and expensive) EHRs allow for decision making assistance
        \vspace{3mm}
        \item Continuing to evolve over time
    \end{itemize}
\end{itemize}

@@

\begin{frame}[fragile]{HIT: Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item Possible cost reduction of hundreds of billions of dollars \\ \scriptsize (Hillestad et al 2005, HA)
\end{itemize}
\end{alertblock}

@@

\note[item]{Why do we care about EHRs or health information technology in the first place?}

\begin{frame}[fragile]{What could this mean for physicians?}
Physicians are the ones who put this technology to use, how they respond will partially determine the technology's impacts

\vspace{3mm}

Immediate Effects:
\begin{itemize}
    \item Large startup costs (time, productivity)
    \item Possible loss of autonomy
\end{itemize}

Long Run Effects:
\begin{itemize}
    \item Productivity could increase
    \item Progressing loss of autonomy as technology progresses
\end{itemize}

\begin{frame}[noframenumbering]{Physician Response to EHRs}
\begin{center}
    \includegraphics[scale=.4]{graphics/News Clip1.PNG}
\end{center}
\end{frame}

@@

Did EHR implementation in hospitals affect physician labor outcomes?
\begin{itemize}
    \item Intensive Margin / Productivity
    \item Extensive Margin
    \item Hospital $\rightarrow$ Office?
\end{itemize}
\end{frame}

@@

\begin{frame}{Literature}
\footnotesize 
\begin{columns}
    \column{0.34\textwidth}
        \centering
        \underline{ EHR $\rightarrow$ Healthcare }
        \begin{itemize}
            \item Small improvements for severe health conditions \\ \vspace{1mm}
            \tiny(Agha 2014, JHE; McCullough et al 2016, RAND; Meyerhoefer et al, ILRR)

@@

@@

            \item Physician burnout: early retirement and less hours worked \\ \vspace{1mm} 
            \tiny (Dewa et al 2014)
        \end{itemize}

\end{columns}
\end{frame}

@@

@@

\begin{frame}{What I Want to Capture}
\begin{itemize}
    \item Initial Implementation
    \vspace{3mm}
    \begin{itemize}
        \item Productivity gains/losses?
        \vspace{3mm}
        \item Retirement or change in place of work?
        \vspace{3mm}
    \end{itemize}
    \item Do the effects change over time after the technology is implemented?

\begin{frame}{About the Data}
    Main measurements come from \underline{CMS Shared Patient Data} (2009-2015)
    \begin{itemize}
        \item Lists 2 NPIs and the number of billings those two entities shared
    \end{itemize}
    \vspace{3mm}

    How I use it:
    \begin{itemize}
        \item Limit the NPIs to physician-hospital pairs, where physicians are PCPs, internists, hospitalists
        \item Limit the billings to those which occur in the same day
        \item Only keep pairs with sufficient number of billings together to suggest a close relationship between physician and hospital
        \item Aggregate to the physician level to avoid spillover effects
    \end{itemize}
\end{frame}

@@

     Hospital patients: 
    \begin{itemize}
        \item Sum of shared patients with all hospitals
    \end{itemize}

@@

     Indicator for switching work setting:
    \begin{itemize}
        \item Positive billings shared with all entities but shared entities with hospitals becomes 0
        \item Database of office-based physicians
    \end{itemize}

@@

Physician-level treatment variable that captures exposure to \textit{any} EHR
\vspace{3mm}
\begin{itemize}
    \item Individual hospital EHR use comes from AHA survey
    \vspace{3mm}
    \item Treatment turns on the first time any of the shared hospitals implements an EHR
\end{itemize}
\end{frame}

@@

\vspace{5mm}

Average size of hospital worked with (measured in beds), average hospital days operating: \underline{AHA Survey}
\end{frame}

\begin{frame}{Summary Statistics}
\centering
    \includegraphics[scale=.8]{Objects/sumstats.pdf}

    \note[item]{Need to divide n by 7 to get actual number of physicians}
\end{frame}

@@

\begin{frame}{Plans}
    \begin{itemize}
        \item Analysis on retirement decision of physicians: requires a better understanding of why physicians would drop out of this Medicaid data
    \end{itemize}
\end{frame}

>>>>

\setbeameroption{slides only}
%\setbeameroption{show notes on second screen=right}
\usepackage{xspace}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}

@@

\begin{frame}{Technology and Displacement}

\begin{frame}[fragile]{What is an Electronic Health Record?}
\begin{itemize}
    \item Technology purchased and implemented by hospitals
\end{itemize}

\begin{itemize}
    \item Used for:
    \vspace{3mm}
    \begin{itemize}
        \item Electronic patient records
        \vspace{3mm}
        \item Decision making assistance
    \end{itemize}
\end{itemize}

@@

\begin{frame}[fragile]{HIT: Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item Possible cost reduction of hundreds of billions of dollars \\ \scriptsize (Hillestad et al 2005)
\end{itemize}
\end{alertblock}

@@

\note[item]{Why do we care about EHRs or health information technology in the first place?}
\note[item]{Verbally explain the bridge to physicians here (the slide I took out)}
\end{frame}

@@

Did EHR implementation in hospitals affect physician labor market decisions?
\begin{itemize}
    \item Intensive Margin / Productivity
    \item Extensive Margin
    \item Hospital $\rightarrow$ Office
\end{itemize}
\end{frame}

@@

\begin{frame}{Literature}
\footnotesize 
\begin{columns}
    \column{0.34\textwidth}
        \centering
        \underline{ EHR $\rightarrow$ Healthcare }
        \begin{itemize}
            \item Small improvements for severe health conditions \\ \vspace{1mm}
            \tiny(Agha 2014; McCullough et al 2016)

@@

            \item No decrease in cost \\ \vspace{1mm}
            \tiny (Agha 2014; Dranove et al 2019)

@@

            \item Physician burnout: early retirement and less hours worked \\ \vspace{1mm} 
            \tiny (Dewa et al 2014)
        \end{itemize}
\end{columns}

\note[item]{\footnotesize Agha 2014 (JHE) uses claims data and difference in difference methodology. Find 1.3 percent increase in billed costs and no improvements in health coutcomes, data is 1998-2005.}

\note[item]{McCullough, Parente, Town 2016, RAND. claims data 2002-2007. Find no reduction in moretality for the median patient, only those who require complex care with coordination.}

\note[item]{Appari, Johnson, Anthony 2013. Health services review. fixed effects linear panel model. quality measure comes from hospital compare, not at the patient level.}

\note[item]{Dranove, Forman, Goldfarb, Greenstein (2019) American Economic Journal: Economic Policy. data from 1996 to 2009. Specific findings: hospitals have a rise in cost after 3 years, some hospitals in certain locations have a decrease in costs after 5 years but others still have a higher cost after 6 years.}

@@

            \item Physician burnout: early retirement and less hours worked \\ \vspace{1mm} 
            \tiny (Dewa et al 2014)
        \end{itemize}
\end{columns}

\note[item]{Zeira and Johnson 2011. }
\end{frame}

@@

\begin{frame}{About the Data}
    Main data: \underline{CMS Shared Patient Data} (2009-2015)
    \begin{itemize}
        \item Lists 2 NPIs and the number of billings those two entities shared
    \end{itemize}
    \vspace{3mm}

    How I use it:
    \begin{itemize}
        \item Physician-hospital pairs (PCPs, internists, hospitalists)
        \item Billings that occur on same day
        \item Close relationship between hospital and physician
        \item Aggregate to the physician level
    \end{itemize}
\end{frame}

@@

     Patients visited in hospitals

@@

     Indicator for switching work setting:
    \begin{itemize}
        \item 0 patients visited but maintaining billable activity
    \end{itemize}

@@

Physician-level treatment variable that captures exposure to \textit{any} EHR
\vspace{3mm}
\begin{itemize}
    \item Individual hospital EHR use comes from AHA survey
    \vspace{3mm}
    \item Treatment turns on the first time any of the physicians' hospitals implements an EHR
\end{itemize}
\end{frame}

@@

\begin{frame}{Data}
\centering
    \includegraphics[scale=.8]{Objects/sumstats.pdf}

@@

\begin{frame}{Plans}
    \begin{itemize}
        \item Analysis on retirement decision of physicians: requires a better understanding of why physicians would drop out of this Medicare data
        \item Investigate other estimation strategies
    \end{itemize}
\end{frame}
